---
title: "HEAL Parks: 2024 Playground Analysis"
author: "Ronald Buie"
date-modified: "`r Sys.Date()`"

format:
  gfm: 
    output-file: "README.md"
    toc: true
    number-sections: true
    fig-height: 12
    fig-width: 16
    prefer-html: true
  pdf:
    toc: true
    number-sections: true
    df-print: kable
    fig-height: 8
    fig-width: 10
  html:
    toc: true
    number-sections: true
    df-print: kable
    fig-height: 12
    fig-width: 16
execute: 
  echo: false
  warning: false
  output: true
editor: 
  markdown: 
    wrap: none
    canonical: true
---

# Front Matter

This document outlines procedures, technical considerations, and analytic results for a 2024 analysis of playgrounds within parks participating in the HEAL Parks Study. The primary purpose of this PDF is technical review by analyst and project managers to confirm the process and data quality.

For general information about the project please review the [git](https://github.com/PHSKC-APDE/HEAL_Parks_Analysis) or contact Seth Schromen-Wawrin.

```{r setup}
# First check if pacman is installed. If not, install it.
if(!"pacman" %in% installed.packages()) {
  install.packages("pacman")
}


# load and install packages using pacman. 
# Pacman will install and load missing packages.
pacman::p_load(doParallel,  #parallalize loops, including geography calculation
               scales, #format ggplot2 axes
               sf, #create spatial objects around long and lat
               kcparcelpop,
               rads, #access apde data
               spatagg, #geospatial maping for populations
               flextable, #generate pretty tables in word
               officer, #generate word docs
               openxlsx2, #read xlsx files
               ggplot2, #create pretty charts
               here,  #simplify calls to local directory
               RCurl, #prepare and parse json/API querries
               jsonlite, #parse json results
               data.table, #data structure and analytics framework used throughout
               kableExtra, #kable extension used to generate pretty output
               utils, #modify directory and file structures
               treemapify) #generate treemaps in ggplot


#configure any package specific options here
options(knitr.table.format = "latex") 

#constants
#CHARTWIDTH <- 3.5 #width for charts to generate for reproduciton
#CHARTHEIGHT <- 2.3 #height of charts to generate for reproduction
RECALCULATEPOPULATIONS <- FALSE
PARALLEL_PROCESS <- TRUE
```

## Major inputs

Inputs to this script are contained at ./inputs/*.* and include

-   A file of park meta data, including address, zip, city, neighborhood, official name, and REDCap name for each park
-   A file of collected SOPARC observations, aggregated to the period level
-   An accompanying file of SOPARC activities

## Output categories

-   tables - tabular outputs of analysis, generally in xlsx format
-   charts - chart outputs of analysis, generally in pdf and/or png format

```{r create target directories}

dir.create("./outputs/charts", recursive = TRUE, showWarnings = FALSE)
dir.create("./outputs/tables", recursive = TRUE, showWarnings = FALSE)
# dir.create("./2024 playground/outputs/data-metadata", recursive = TRUE, showWarnings = FALSE)
```

## This is a quarto generated document

By rendering/knitting the qmd file, the analysis is re-executed, this document rebuilt, and new outputs are generated. To learn more about Quarto see \[https://quarto.org\].

# Setup & Environment

This script was last executed using `r version$version.string`.

```{r define helpful functions}

integer_breaks <- function(n = 5, ...) {
  #accespt none
  #returns break point
  #used to get integers on short axies in ggplot
  # by Joshua Cook: https://www.r-bloggers.com/2019/11/setting-axes-to-integer-values-in-ggplot2/

  fxn <- function(x) {
  breaks <- floor(pretty(x, n, ...))
  names(breaks) <- attr(breaks, "labels")
  breaks
  }
return(fxn)
}

```

# Data Preperation

This analysis relies on park observations and park activity tables created in the annual SOPARC study

```{r import data}
ParkObservations <- fread("./inputs/SOPARCAnalysisSetAggregatedPeriods.csv")



ParksActivities <- fread("./inputs/SOPARCActivitiesExpanded.csv")


ParkSupplamentalInfo <- read_xlsx("./inputs/Parks_MASTER.xlsx", sheet = "2024Supplemental")
ParkSupplamentalInfo[ParkSupplamentalInfo$Restroom == "Yes ",]$Restroom <- "Yes"
ParkSupplamentalInfo[ParkSupplamentalInfo$`Primary_Age_Group_2-5` == "Yes ",]$`Primary_Age_Group_2-5` <- "Yes"


ParkFacilities <- read_xlsx("./inputs/Parks_MASTER.xlsx", sheet = "ParkFacility")


ParkMetadata <- as.data.table(read_xlsx("./inputs/Parks_MASTER.xlsx", sheet = "ParkList"))


```

```{r merged tables}
#these are tables used in most/all analyses

ParkFull <- as.data.table(merge(ParkObservations, ParkSupplamentalInfo, by.x = "park_name_full", by.y = "Park Name"))


```

```{r merging facility info}


ParkFull <- merge(ParkFull, ParkFacilities, by.x = c("park_name_full", "tar_area"), by.y = c("Park Name", "Scan Area"))



```

```{r keeping on parks with playgrounds}

for(park in unique(ParkFull$park_name_full)) {
  if(nrow(ParkFull[park_name_full == park & playground == 1,]) == 0) {
    ParkFull <- ParkFull[park_name_full != park,]
  }
}


```

## retrieving child populations within .5 mile radius

```{r calculate population of children}
#this chunk is very slow, so if the data already exists, do not rerun.
skip <- FALSE
if(file.exists("./inputs/Child Park Populations.csv")){
  ChildParkPopulationTable <- fread("./inputs/Child Park Populations.csv")
  if("childPopulationHalfMile" %in% names(ChildParkPopulationTable) &
     all(unique(ParkFull$park_name_full) %in% ChildParkPopulationTable[censusYear == 2022,]$park_name_full)) {
    skip <- TRUE #yay no need to calculate 
  }
}
if(!skip | RECALCULATEPOPULATIONS) {
  
  
  
  crsString <- "EPSG:2926" #preferred coordinate reference system for WA
  
  ParkGeos <- unique.data.frame(ParkFull[,.("park" = park_name_full, "long" = Longitude, "lat" = Latitude)]) #create a prototype data frame for holding our geometry. This has the raw lat long data
  
  ParkGeos <- st_as_sf(x = ParkGeos,coords = c("long","lat"), crs = "EPSG:4326") #turn teh raw data into a geometry using crs 2926 (this is one recomended for WA)
  ParkGeos <- st_transform(ParkGeos, st_crs(crsString))
  ParkGeos <- st_buffer(ParkGeos, units::set_units(0.5, mile)) # units::set_units is intelligent about types of units so you can specify "mile". st_buffer will create a perimeter of the provided units.
  
  BlockShapes <- st_read("//dphcifs/APDE-CDIP/Shapefiles/Census_2020/block/kc_block.shp") 
  BlockShapes <- st_transform(BlockShapes, st_crs(crsString)) #conform to same crs
  
  # ggplot() + geom_sf(data = BlockShapes, fill = NA) +
  #   geom_sf(data = ParkGeos, fill = NA, color = 'purple')
  
  #get KC population estimates. Note, if a radius desired extends outside the county, then need to expand. Will greatly slow down calculations
  #if not saved, pull and save. Otherwise, load from save
  if(file.exists("./inputs/KCChildPops.csv")) {
    KCPops <- read.csv("./inputs/KCChildPops.csv")
  } else {
    KCPops <- get_population(geo_type = "blk", kingco = T, year = 2022, ages = 0:14)
    write.csv(x = KCPops, "./inputs/KCChildPops.csv", row.names = F)
  }
  
  #Create data frame for final results
  ChildParkPopulationTable <- data.table("park_name_full" = unique(ParkFull$park_name_full), "censusYear" = 2022, "childPopulationHalfMile" = 0)
  
#  if(PARALLEL_PROCESS){
#  numCores <- detectCores()-1
#  cl <- makeCluster(numCores, outfile = "")
#  registerDoParallel(cl)
#  ChildParkPopulationTable[parkNameFull == ParkGeos[rowIndex,]$park, ]$childPopulationHalfMile <- foreach (rowIndex = 1:nrow(ParkGeos), #.combine=rbind) %dopar% {
#    CW <- create_xwalk(BlockShapes, ParkGeos[rowIndex,], "GEOID20", "park",min_overlap = 0.00001)
#    CWPop <- merge(CW, KCPops, by.x = "source_id", by.y = "geo_id")
#    weightedPop <- sum(CWPop$s2t_fraction * CWPop$pop)
#    weightedPop 
#  }
#  stopCluster(cl)
#} else {
  #loop through parks and generate crosswalk for each park.
  for(rowIndex in 1:nrow(ParkGeos)) {
    CW <- create_xwalk(BlockShapes, ParkGeos[rowIndex,], "GEOID20", "park",min_overlap = 0.00001)
    CWPop <- merge(CW, KCPops, by.x = "source_id", by.y = "geo_id")
    weightedPop <- sum(CWPop$s2t_fraction * CWPop$pop)
    
    ChildParkPopulationTable[park_name_full == ParkGeos[rowIndex,]$park, ]$childPopulationHalfMile <- weightedPop 
    #print(paste0("Completed #", rowIndex," (",ParkGeos[rowIndex,park],")"))
    print(paste0("Completed ", rowIndex, " of ", nrow(ParkGeos)))
  }
#}
  


  write.csv(ChildParkPopulationTable , "./outputs/Child Park Populations.csv", row.names = F)
  write.csv(ChildParkPopulationTable ,"./inputs/Child Park Populations.csv", row.names = F)
}

  
ParkFull <- merge(ParkFull, ChildParkPopulationTable, by = "park_name_full")

```

# Analyses

Notes:

-   Analyses of "average park use" aggregate all playground within the park (effectively calculated as one large playground)
-   Multiple park characteristics in the supplamental information file have n/a or tbd for parks that, including for parks that have playgrounds. Depending on the analysis, these may show up as NA, tbd, or 0

### How does the observed use of playgrounds differ between parks located in opportunity areas and parks not located in opportunity areas?

## Chart of Park Catchment, sorted by Opportunity Area

```{r}
DT <- copy(ParkFull)

DTplot <- unique(DT[, Playground_Age := (2024-as.numeric(DT$Playground_Age)) +1][, .(park_name_full, populationHalfMile, Opportunity_Area)])

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = populationHalfMile)) +
  facet_wrap(~Opportunity_Area, scales = "free_y") +
  xlab("Population Within 1/2 Mile") +
  ylab("Park Name") +
  ggtitle("Catchment Given Opportunity Area Status") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)



```

## Chart of Youth Park Catchment, sorted by Opportunity Area

```{r}

DT <- copy(ParkFull)

DTplot <- unique(DT[, Playground_Age := (2024-as.numeric(DT$Playground_Age)) +1][, .(park_name_full, childPopulationHalfMile, Opportunity_Area)])

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = childPopulationHalfMile)) +
  facet_wrap(~Opportunity_Area, scales = "free_y") +
  xlab("0-14 Population Within 1/2 Mile") +
  ylab("Park Name") +
  ggtitle("0-14 Year Old Catchment Given Opportunity Area Status") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)

```

### Chart of Average Playground Use, sorted by Opportunity Area

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Opportunity_Area, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( Opportunity_Area, park_name_full)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = avg_Daily_Users)) +
  facet_wrap(~Opportunity_Area, scales = "free_y") +
  xlab("Average Daily number of playground users") +
  ylab("Park Name") +
  ggtitle("Average Daily Number of Playground Users Given \n Opportunity Area Status") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

### Chart of Peak Playground Use, sorted by Opportunity Area

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(max_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Opportunity_Area, park_name_full)][,lapply(.SD, max, na.rm = TRUE), .SDcols = c("max_Daily_Users") ,by = .( Opportunity_Area, park_name_full)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = max_Daily_Users)) +
  facet_wrap(~Opportunity_Area, scales = "free_y") +
  xlab("Max Daily number of playground users") +
  ylab("Park Name") +
  ggtitle("Max Daily number of Playground Users Given \n Opportunity Area Status") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

## How does the observed use of playgrounds differ between parks of similar park categories?

### Chart of Park Catchment, sorted by Park Category

```{r}
DT <- copy(ParkFull)

DTplot <- unique(DT[, Playground_Age := (2024-as.numeric(DT$Playground_Age)) +1][, .(park_name_full, populationHalfMile, Park_Category_Local)])

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = populationHalfMile)) +
  facet_wrap(~Park_Category_Local, scales = "free_y") +
  xlab("Population Within 1/2 Mile") +
  ylab("Park Name") +
  ggtitle("Catchment Given Park Category") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)
```

### Chart of Youth Park Catchment, sorted by Park Category

```{r}

DT <- copy(ParkFull)

DTplot <- unique(DT[, Playground_Age := (2024-as.numeric(DT$Playground_Age)) +1][, .(park_name_full, childPopulationHalfMile, Park_Category_Local)])

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = childPopulationHalfMile)) +
  facet_wrap(~Park_Category_Local, scales = "free_y") +
  xlab("0-14 Population Within 1/2 Mile") +
  ylab("Park Name") +
  ggtitle("0-14 Year Old Catchment Given Park Category") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)

```

### Chart of Average Playground Use, sorted by Park Category

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Park_Category_Local, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( Park_Category_Local, park_name_full)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = avg_Daily_Users)) +
  facet_wrap(~Park_Category_Local, scales = "free_y") +
  xlab("Average Daily number of playground users") +
  ylab("Park Name") +
  ggtitle("Average Daily Number of Playground Users Given \n Park Category") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

### Chart of Peak Playground Use, sorted by Park Category

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(max_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Park_Category_Local, park_name_full)][,lapply(.SD, max, na.rm = TRUE), .SDcols = c("max_Daily_Users") ,by = .( Park_Category_Local, park_name_full)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = max_Daily_Users)) +
  facet_wrap(~Park_Category_Local, scales = "free_y") +
  xlab("Max Daily number of playground users") +
  ylab("Park Name") +
  ggtitle("Max Daily number of Playground Users Given \n Park Category") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

### Chart of Average Playground Popularity, sorted by Park Category

Average Popularity is the average of the daily popularity rate, where popularity rate is the proportion of park users (in a day) who were in the playground

```{r}

DT <- copy(ParkFull[occupied == "Yes"])

DTplot <- DT[, .(users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, park_name_full, Park_Category_Local, playground)][, .(users = sum(users)),by = .(park_name_full,  Park_Category_Local, playground)][,allusers := sum(users), by = .(park_name_full, Park_Category_Local)][playground == 1, popularity := users/allusers][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("popularity") ,by = .(park_name_full, Park_Category_Local)]


ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = popularity)) +
  facet_wrap(~Park_Category_Local, scales = "free_y") +
  xlab("Average Playground Popularity Rate") +
  ylab("Park Name") +
  ggtitle("Playground Popularity Given Category") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)
```

## How does the observed use of playgrounds differ between playgrounds with different complexities of playgrounds?

### Chart of Average Playground Use, sorted by Play Elements

Notes:

-   Garfield is missing a value

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Elements, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( Elements, park_name_full)][, Elements := as.numeric(Elements)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(x = Elements, fill = avg_Daily_Users, y = reorder(park_name_full, Elements))) +
  xlab("Number of Elements") +
  ylab("Park Name") +
  guides(fill=guide_legend(title="Average Daily Users")) +
  ggtitle("Average Daily Number of Playground Users Given \n Number of Elements") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

### Chart of Peak Playground Use, sorted by Play Elements

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(max_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Elements, park_name_full)][,lapply(.SD, max, na.rm = TRUE), .SDcols = c("max_Daily_Users") ,by = .( Elements, park_name_full)][, Elements := as.numeric(Elements)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(x = Elements, fill = max_Daily_Users, y = reorder(park_name_full, Elements))) +
  xlab("Number of Elements") +
  ylab("Park Name") +
  guides(fill=guide_legend(title="Max Daily Users")) +
  ggtitle("Max Daily Number of Playground Users Given \n Number of Elements") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)
```

### Chart of Average Playground Popularity, sorted by Play Elements

```{r}

DT <- copy(ParkFull[occupied == "Yes"])

DTplot <- DT[, .(users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, park_name_full, Elements, playground)][, .(users = sum(users)),by = .(park_name_full,  Elements, playground)][,allusers := sum(users), by = .(park_name_full, Elements)][playground == 1, popularity := users/allusers][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("popularity") ,by = .(park_name_full, Elements)][, Elements := as.numeric(Elements)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(x = Elements, fill = popularity, y = reorder(park_name_full, Elements))) +
  xlab("Number of Elements") +
  ylab("Park Name") +
  guides(fill=guide_legend(title="Average Popularity Rate")) +
  ggtitle("Playground Popularity Given \n Number of Elements") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)
```

## How does the observed use of playgrounds compare to the age of the playground?

### Chart of Average Playground Use, sorted by Playground Installation Date

garfield is missing data

```{r}
DT <- copy(ParkFull)

DTplot <- DT[, Playground_Age := (2024-as.numeric(DT$Playground_Age)) +1][playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Playground_Age, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( Playground_Age, park_name_full)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(x = as.numeric(Playground_Age), fill = avg_Daily_Users, y = reorder(park_name_full, Playground_Age))) +
  xlab("Years Age of Playground (1 = Built in 2024)") +
  ylab("Park Name") +
  guides(fill=guide_legend(title="Average Daily Users")) +
  ggtitle("Average Daily Number of Playground Users Given \n Playground Age") +
  scale_y_discrete(limits = rev) #+
  scale_x_continuous(labels = scales::comma)


```

### Chart of Peak Playground Use, sorted by Playground Installation Date

```{r}
DT <- copy(ParkFull)

DTplot <- DT[, Playground_Age := (2024-as.numeric(DT$Playground_Age)) +1][playground == 1, .(max_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Playground_Age, park_name_full)][,lapply(.SD, max, na.rm = TRUE), .SDcols = c("max_Daily_Users") ,by = .( Playground_Age, park_name_full)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(x = as.numeric(Playground_Age), fill = max_Daily_Users, y = reorder(park_name_full, Playground_Age))) +
  xlab("Years Age of Playground (1 = Built in 2024)") +
  ylab("Park Name") +
  guides(fill=guide_legend(title="Max Daily Users")) +
  ggtitle("Max Number of Playground Users Given \n Playground Age") +
  scale_y_discrete(limits = rev) #+
  scale_x_continuous(labels = scales::comma)


```

### Chart of Playground Installation Date and Opportunity Area (denoting differences in Park Category)

```{r}
DT <- copy(ParkFull)

DTplot <- unique(DT[, Playground_Age := (2024-as.numeric(DT$Playground_Age)) +1][, .(park_name_full, Playground_Age, Opportunity_Area)])

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = Playground_Age)) +
  facet_wrap(~Opportunity_Area, scales = "free_y") +
  xlab("Years Age of Playground (1 = Built in 2024)") +
  ylab("Park Name") +
  ggtitle("Playground Age Given Opportunity Area Status") +
  scale_y_discrete(limits = rev)



```

## How does the observed use of playgrounds differ when a park includes certain features?

### Chart of Average Playground Use, based on if park contains a restroom, separated by park category

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Park_Category_Local, Restroom, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( Park_Category_Local, Restroom,park_name_full)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = avg_Daily_Users)) +
  facet_grid(Restroom~Park_Category_Local, scales = "free_y") +
  xlab("Average Daily number of playground users") +
  ylab("Park Name") +
  ggtitle("Average Daily Number of Playground Users Given \n Category and Restroom") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

### Chart of Average Playground Use, based on if park contains a picnic area, separated by park category

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Park_Category_Local, PicnicArea, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( Park_Category_Local, PicnicArea,park_name_full)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = avg_Daily_Users)) +
  facet_grid(PicnicArea~Park_Category_Local, scales = "free_y") +
  xlab("Average Daily number of playground users") +
  ylab("Park Name") +
  ggtitle("Average Daily Number of Playground Users Given \n Category and Picnic Area") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

### Chart of Average Playground Use, based on if park contains a sports field, separated by park category

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Park_Category_Local, SportsField, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( Park_Category_Local, SportsField, park_name_full)]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = avg_Daily_Users)) +
  facet_grid(SportsField~Park_Category_Local, scales = "free_y") +
  xlab("Average Daily number of playground users") +
  ylab("Park Name") +
  ggtitle("Average Daily Number of Playground Users Given \n Category and SportsField") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

## How does the age of observed playground users differ between park groups (size, feature qualities, etc.)?

### Chart of Average Playground Use, sorted by Play Element, separated by Primary Age Group

garfield is missing, not actually "no"

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Elements, `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( Elements,  `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`, park_name_full)][, Elements := as.numeric(Elements)][, `Primary_Age_Group_2-5` := ifelse(`Primary_Age_Group_2-5` == "Yes", "2-5", "No")][, `Primary_Age_Group_5-12` := ifelse(`Primary_Age_Group_5-12` == "Yes", "5-12", "No")]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = avg_Daily_Users, fill = Elements)) +
  facet_wrap(~`Primary_Age_Group_2-5` + `Primary_Age_Group_5-12`, scales = "free_y") +
  xlab("Average Daily number of playground users") +
  ylab("Park Name") +
  ggtitle("Average Daily Number of Playground Users Given \n Primary Age Group and Elements") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

### Chart of Peak Playground Use, sorted by Play Element, separated by Primary Age Group

garfield is missing, not actually "no"

```{r}
DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(max_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Elements, `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`, park_name_full)][,lapply(.SD, max, na.rm = TRUE), .SDcols = c("max_Daily_Users") ,by = .( Elements,  `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`, park_name_full)][, Elements := as.numeric(Elements)][, `Primary_Age_Group_2-5` := ifelse(`Primary_Age_Group_2-5` == "Yes", "2-5", "No")][, `Primary_Age_Group_5-12` := ifelse(`Primary_Age_Group_5-12` == "Yes", "5-12", "No")]

ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = max_Daily_Users, fill = Elements)) +
  facet_wrap(~`Primary_Age_Group_2-5` + `Primary_Age_Group_5-12`, scales = "free_y") +
  xlab("Max Daily number of playground users") +
  ylab("Park Name") +
  ggtitle("Max Daily Number of Playground Users Given \n Primary Age Group and Elements") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)


```

### Chart of Playground Popularity, sorted by Play Element, separated by Primary Age Group

```{r}

DT <- copy(ParkFull[occupied == "Yes"])

DTplot <- DT[, .(users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, Elements, park_name_full,  `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`, playground)][, .(users = sum(users)),by = .(Elements, park_name_full,  `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`, playground)][,allusers := sum(users), by = .( Elements, park_name_full,  `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`)][playground == 1, popularity := users/allusers][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("popularity") ,by = .(Elements, park_name_full, `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`)][, Elements := as.numeric(Elements)][, `Primary_Age_Group_2-5` := ifelse(`Primary_Age_Group_2-5` == "Yes", "2-5", "No")][, `Primary_Age_Group_5-12` := ifelse(`Primary_Age_Group_5-12` == "Yes", "5-12", "No")]


ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = popularity, fill = Elements)) +
  facet_wrap(~`Primary_Age_Group_2-5` + `Primary_Age_Group_5-12`, scales = "free_y") +
  xlab("Popularity") +
  ylab("Park Name") +
  ggtitle("Playground Popularity Given \n Primary Age Group and Elements") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)

```

### Chart of Playground Popularity specifically of observed child and teen, sorted by Play Element, separated by Primary Age Group

Here, popularity is parameterized as

$\frac{\text{num of children and teens in playground}} {\text{num of children and teens in park}}$

```{r}

DT <- copy(ParkFull[occupied == "Yes"])

DTplot <- DT[, .(users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, na.rm = TRUE)), by = .(datePOSIX, Elements, park_name_full,  `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`, playground)][, .(users = sum(users)),by = .(Elements, park_name_full,  `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`, playground)][,allusers := sum(users), by = .( Elements, park_name_full,  `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`)][playground == 1, popularity := users/allusers][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("popularity") ,by = .(Elements, park_name_full, `Primary_Age_Group_2-5`, `Primary_Age_Group_5-12`)][, Elements := as.numeric(Elements)][, `Primary_Age_Group_2-5` := ifelse(`Primary_Age_Group_2-5` == "Yes", "2-5", "No")][, `Primary_Age_Group_5-12` := ifelse(`Primary_Age_Group_5-12` == "Yes", "5-12", "No")]


ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(y = park_name_full, x = popularity, fill = Elements)) +
  facet_wrap(~`Primary_Age_Group_2-5` + `Primary_Age_Group_5-12`, scales = "free_y") +
  xlab("Popularity") +
  ylab("Park Name") +
  ggtitle("Playground Popularity Among Children Given \n Primary Age Group and Elements") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(labels = scales::comma)

```

## How accessible is the park by transportation?

## By walking: Chart of Playground Use by Walkscore of park

```{r}

DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, WalkScore, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( WalkScore, park_name_full)]


ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(x = as.numeric(WalkScore), fill = avg_Daily_Users, y = reorder(park_name_full, WalkScore))) +
  xlab("Walkability Score") +
  ylab("Park Name") +
  guides(fill=guide_legend(title="Average Daily Users")) +
  ggtitle("Average Daily Number of Playground Users Given \n Playground Age") +
  scale_y_discrete(limits = rev) #+
  scale_x_continuous(labels = scales::comma)


```

## By transit: Chart of Playground Use by Transitscore of park

```{r}

DT <- copy(ParkFull)

DTplot <- DT[playground == 1, .(avg_Daily_Users = sum(num_child_prim, num_child_snd, num_child_spec, num_teen_prim, num_teen_snd, num_teen_spec, num_adult_prim, num_adult_snd, num_adult_spec, num_senior_prim, num_senior_snd, num_senior_prim, na.rm = TRUE)), by = .(datePOSIX, TransitScore, park_name_full)][,lapply(.SD, mean, na.rm = TRUE), .SDcols = c("avg_Daily_Users") ,by = .( TransitScore, park_name_full)][, TransitScore := as.numeric(TransitScore)]


ggplot(DTplot) +
  geom_bar(stat = "identity", position = "stack", aes(x = as.numeric(TransitScore), fill = avg_Daily_Users, y = reorder(park_name_full, TransitScore))) +
  xlab("Transit Score") +
  ylab("Park Name") +
  guides(fill=guide_legend(title="Average Daily Users")) +
  ggtitle("Average Daily Number of Playground Users Given \n Playground Age") +
  scale_y_discrete(limits = rev) #+
  scale_x_continuous(labels = scales::comma)


```
